Exploring clinical and non-clinical(food) Listeria monocytogenes isolates 

### 2025-04-06 code/analysis-code/statistical-analysis.qmd
adding some figures to manuscript

ggplot(mydata, aes(x = days_since_start, y = clinical_numeric)) +
  geom_jitter(height = 0.05, width = 5, alpha = 0.5) +  
  geom_line(data = new_data, aes(x = days_since_start, y = predicted_prob), color = "blue", size = 1) +
  labs(title = "Logistic Regression: Probability of Clinical Isolates Over Time",
       x = "Days Since Start",
       y = "Probability of Outcome (Yes = 1)") +
  theme_minimal()

```
```{r}
ggplot(mydata, aes(x = days_since_start, fill = factor(clinical_numeric))) +  # Fill by outcome (0 or 1)
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 30, 
                 color = "black") +  # Standard histogram
  labs(title = "Histogram of Outcomes Over Time",
       x = "Days Since Start",
       y = "Count of Outcome (0 or 1)",
       fill = "Outcome") +
  theme_minimal()

>>>>

ggplot(mydata, aes(x = days_since_start, y = clinical_numeric)) +
  geom_jitter(height = 0.05, width = 5, alpha = 0.5) +  
  geom_line(data = mydata, aes(x = days_since_start, y = predicted_prob), color = "blue", size = 1) +
  labs(title = "Logistic Regression: Probability of Clinical Isolates Over Time",
       x = "Days Since Start",
       y = "Probability of Outcome (Yes = 1)") +
  theme_minimal()

```
```{r}
p <- ggplot(mydata, aes(x = days_since_start, fill = factor(clinical_numeric))) +  # Fill by outcome (0 or 1)
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 30, 
                 color = "black") +  # Standard histogram
  labs(title = "Histogram of clinical isolates Over Time",
       x = "Days Since Start",
       y = "Count of Outcome (0 or 1)",
       fill = "Outcome") +
  theme_minimal()

figure_file=here("results/figures","clinicalxtime.png")
ggsave(filename=figure_file, plot = p, width = 8, height=6)
```

### 2025-04-06 code/eda-code/eda.qmd
adding some figures to manuscript

#Making a table for the clinical variable 
```{r}
table(processeddata$clinical)
```
#doing the same for food
```{r}
table(processeddata$food)
```
#Now lets visualize these tables using bar graphs
```{r}
ggplot(mydata, aes(x = clinical)) + 
  geom_bar(fill = "steelblue") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()

@@

ggplot(mydata, aes(x = food)) + 
  geom_bar(fill = "orange") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()

>>>>

#Making a table for the clinical variable 
```{r}
table(processeddata$clinical)
```
#doing the same for food
```{r}
table(processeddata$food)
```
#Now lets visualize these tables using bar graphs
```{r}
bar1 <- ggplot(mydata, aes(x = clinical)) + 
  geom_bar(fill = "steelblue") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()

@@

p1 <- ggplot(mydata, aes(x = food)) + 
  geom_bar(fill = "orange") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()
figure_file=here("results/figures","foodisolatecount.png")
ggsave(filename=figure_file, plot = p1, width = 8, height=6)

### 2025-04-06 products/manuscript/Manuscript.qmd
adding some figures to manuscript

\
This study aims to investigate trends in listeria isolates found in clinical and non-clinical samples in efforts to predict cases.  This study will investigate time of a year and number of non-clinical isolates as a predictor for clinical isolates, and food type as a predictor for food isolates , non-clinical isolates that originate in food products. The data source being used to conduct this study is the National Center for Biotechnology Information\'s Pathogen Detection browser. Pathogen Detection is a culmination of pathogen genomic sequences from federal, state, private, and foreign laboratories. It is utilized constantly by public health officials,epidemiologists, and scientists to identify chains of transmission. The new isolates are analyzed in real time to aid in that identification by developing clusters. Listeria monocytogenes is a nationally notifiable disease so, many laboratories opt to share their data nationally through the Pathogen Detection Browser. By utilizing the most up to date data and identifying trends, this study hopes to provide information to form new guidelines on protecting the public from Listeriosis infection.

@@

##

```         
```

## Data import and cleaning

@@

## Statistical analysis

*Explain anything related to your statistical analyses.*

@@

The full code for the EDA is under code-\>eda-code-\>eda.qmd

@@

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

@@

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

@@

*Listeria monocytogenes* is a ubiquitous, gram positive, intra-cellular pathogen renowned for its adaptability and resilience (cite). This bacterium can be found in a wide range of environments such as soil, water, animals, and vegetation. Due to its ability to proliferate in food products and low temperatures, it is considered a major food-borne illness. The ingestion of Listeria-infected food may lead to listeriosis, a gastrointestinal illness that causes vomiting, diarrhea, and even death. It is of even greater concern in pregnant individuals, as its forceful invasion tactics allow the bacterium to cross the placental barrier, infecting the fetus; this can lead to miscarriages and stillbirths\[cite\].Listeriosis causes 1,600 infections and 260 deaths each year with an over 95% hospitalization rate. Listeria can be found in a wide range of meats, produce, and dairy, such as deli meats, soft cheese, and melons. The Centers for Disease Control and Prevention has deemed *Listeria monocytogenes* as a nationally notifiable disease which is a disease that federal agencies, healthcare providers, and laboratories are required to report. All positive isolate data found in the environment, food samples, clinical samples and animal samples are uploaded to the National Center for Biotechnology Information's Pathogen Detection system.

>>>>

\
This study aims to investigate trends in listeria isolates found in clinical and non-clinical samples in efforts to predict cases.  This study will investigate time of a year and number of non-clinical isolates as a predictor for clinical isolates, and food type as a predictor for food isolates , non-clinical isolates that originate in food products. The data source being used to conduct this study is the National Center for Biotechnology Information's Pathogen Detection browser. Pathogen Detection is a culmination of pathogen genomic sequences from federal, state, private, and foreign laboratories. It is utilized constantly by public health officials,epidemiologists, and scientists to identify chains of transmission. The new isolates are analyzed in real time to aid in that identification by developing clusters. Listeria monocytogenes is a nationally notifiable disease so, many laboratories opt to share their data nationally through the Pathogen Detection Browser. By utilizing the most up to date data and identifying trends, this study hopes to provide information to form new guidelines on protecting the public from Listeriosis infection.

@@

@@

@@

```{r}
knitr::include_graphics(here("results/figures", "foodisolatecount.png"))
```

@@

@@

#Need to insert figures

```{r}
knitr::include_graphics(here("results/figures", "clinicalxtime.png"))

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."...................;;;;;;;;;;;;;;;::::::::::::::::P
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

@@
